---
title: "Bayesian Model Selection and Prediction"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  word_document: default
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(statsr)
library(BAS)
library(knitr)
```

### Load data

```{r}
load("movies.Rdata")

d1<-as.data.frame(movies)

dim(d1)

```



* * *

## Part 1: Data

The data set is a randomly selected group of 651 movies produced and released between 1970 and 2014.  Since this data is observational and not related to an expirement with random assignment, we can not look for causal relationships.  However, since the movies were randomly selected, we can look for associations and apply them to the population of interest, movies produced and released for consumption in the US market.

There is little information on how the movies were randomly selected.  One might want to check to see if the composition of the the factor varialbes in the sample is representative of the total population of movies released within the date range.  There is also a potential for bias in the audience scores if we find that the populaton providing their opinions to IMDb and Rotten Tomatoes is not a good representation of the population viewing movies in the US market.


* * *

## Part 2: Data manipulation

First, I'll select the variables I'll need from the original data set, and use the mutate() function to create the new factors required for the model.

```{r}
d2<-d1 %>% 
  select(title_type, genre, runtime, mpaa_rating, thtr_rel_year,
         thtr_rel_month, imdb_rating, imdb_num_votes, critics_score,
         best_pic_nom, best_pic_win, best_actor_win, best_actress_win,
         best_dir_win, top200_box, audience_score) %>%
  mutate(feature_film=ifelse(title_type=="Feature Film", "yes", "no")) %>%
  mutate(drama=ifelse(genre=="Drama", "yes", "no")) %>%
  mutate(mpaa_rating_R=ifelse(mpaa_rating=="R", "yes", "no")) %>%
  mutate(oscar_season=ifelse(thtr_rel_month>9, "yes", "no")) %>%
  mutate(summer_season=ifelse(thtr_rel_month>4 & thtr_rel_month<9, "yes", "no"))

```

Next, I'll select only the varialbes specified for EDA and Modelling, putting the dependant variable first.

```{r}
movdat<-d2 %>%
  select(audience_score, feature_film, drama, runtime, mpaa_rating_R, thtr_rel_year,
         oscar_season, summer_season, imdb_rating, imdb_num_votes,
         critics_score, best_pic_nom, best_pic_win, best_actor_win,
         best_actress_win, best_dir_win, top200_box)
```

Finally, I will check for any observations with missing data and remove any that I find.  Then I'll look at the dimensions of the final data set.

```{r}
cc<-complete.cases(movdat)
movdat<-movdat[cc,]
na<-sum(is.na(movdat))
na
dim(movdat)
```

We see that only one record is removed and that the final data set has 650 observations of 17 variables.

* * *

## Part 3: Exploratory data analysis

Now that we have our data set, let's take a look relationships between the dependant variable (audience_score) and the numeric predictors.

First, we'll look at some summary statistics:

```{r fig.align='center'} 

edan<-select_if(movdat, is.numeric)

s1<-summary(edan$audience_score)
s2<-summary(edan$runtime)
s3<-summary(edan$thtr_rel_year)
s4<-summary(edan$imdb_rating)
s5<-summary(edan$imdb_num_votes)
s6<-summary(edan$critics_score)

stats<-names(edan)

sumtable<-as.data.frame(rbind(s1,s2,s3,s4,s5,s6))
sumtable<-as.data.frame(cbind(stats, sumtable))

sumtable<-sumtable %>%
  mutate(Skew=ifelse(Mean>Median, "Right", "Left"))

kable(sumtable)

```

We see from the summary statistics that the IMDb number of votes is exrremely right skewed.  Most of the films are between and hour and a half and two hours, with a few much longer.  The IMDb ratings and Rotten Tomatoes scores are left skewed with the ratings/scores bunching up a bit at the higher end of the range.

The following code will plot the variables from the table above in a correlation matrix.  This will allow us to see how these variables correlate with the audience_score variable, and help us look for colinearity between independent variables. It also provides a density curve which will help us look at the distributions.

```{r fig.align='center',fig.width=11, fig.height=8}
p1<-ggpairs(edan)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))

p1

```

From these plots we see that the IMDb rating and critics score are both highly correlated with our dependant variable, audience score.  They are also highly correlated with each other.  It will be interesting to see how the bayesian model seletion process handles the colinearity between the IMDb rating and the critics score.

Now we'll evaluate the new factor varialbes we created.  We'll use summary tables, box plots and density plots.  First up is the feature_film variable.

#### Feature Film

```{r fig.align='center', fig.width=11, fig.height=6}

eda<-movdat %>%
  select(audience_score, feature_film, drama,
         mpaa_rating_R, oscar_season, summer_season)

eda1<-eda %>%
  group_by(feature_film) %>%
  summarise(n(), min(audience_score), quantile(audience_score, .25), 
            median(audience_score), mean(audience_score),
            quantile(audience_score, .75), max(audience_score))

names(eda1)<-c("FeatureFilm", "Count", "Min", "1st Qu.", "Median", "Mean",
               "3rd Qu.", "Max")
eda1<-mutate(eda1, Skew=ifelse(Mean>Median, "Right", "Left"))

p2<-ggplot(eda, aes(x=factor(feature_film), y=audience_score, fill=factor(feature_film)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Box Plot")+
  labs(x="Feature Film?", y="Audience Score")

mu2<-eda %>%
  group_by(feature_film) %>%
  summarise(Mean=mean(audience_score))

p2a<-ggplot(eda, aes(x=audience_score, color=feature_film, fill=feature_film))+
  geom_density()+facet_grid(feature_film~.)+
  geom_vline(data=mu2, aes(xintercept=Mean),
             linetype="dashed")+
  theme(legend.position = "right", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Distribution")+
  labs(x="Audience Score", y="Density")


```

```{r echo=F, fig.align='center', fig.width=11, fig.height=4}

kable(eda1)

grid.arrange(p2, p2a, ncol=2)

```

We see that there is a significant difference in the mean audience scores between feature films and non-feature films.  There are far fewer non-feature films, but they tend to have higher audience scores.

We'll  take a similar look at the other factor variables we created.

#### Drama

```{r echo=F}

eda2<-eda %>%
  group_by(drama) %>%
  summarise(n(), min(audience_score), quantile(audience_score, .25), 
            median(audience_score), mean(audience_score),
            quantile(audience_score, .75), max(audience_score))

names(eda2)<-c("Drama", "Count", "Min", "1st Qu.", "Median", "Mean",
               "3rd Qu.", "Max")
eda2<-mutate(eda2, Skew=ifelse(Mean>Median, "Right", "Left"))


p3<-ggplot(eda, aes(x=factor(drama), y=audience_score, fill=factor(drama)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Box Plot")+
  labs(x="Drama?", y="Audience Score")

mu3<-eda %>%
  group_by(drama) %>%
  summarise(Mean=mean(audience_score))

p3a<-ggplot(eda, aes(x=audience_score, color=drama, fill=drama))+
  geom_density()+facet_grid(drama~.)+
  geom_vline(data=mu3, aes(xintercept=Mean),
             linetype="dashed")+
  theme(legend.position = "right", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Distribution")+
  labs(x="Audience Score", y="Density")
```

```{r echo=F, fig.align='center', fig.width=11, fig.height=4}

kable(eda2)

grid.arrange(p3, p3a, ncol=2)

```

Not much here....

#### R Rating

```{r echo=F}

eda3<-eda %>%
  group_by(mpaa_rating_R) %>%
  summarise(n(), min(audience_score), quantile(audience_score, .25), 
            median(audience_score), mean(audience_score),
            quantile(audience_score, .75), max(audience_score))

names(eda3)<-c("R Rating", "Count", "Min", "1st Qu.", "Median", "Mean",
               "3rd Qu.", "Max")
eda3<-mutate(eda3, Skew=ifelse(Mean>Median, "Right", "Left"))

p4<-ggplot(eda, aes(x=factor(mpaa_rating_R), y=audience_score, fill=factor(mpaa_rating_R)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Box Plot")+
  labs(x="Rated R?", y="Audience Score")

mu4<-eda %>%
  group_by(mpaa_rating_R) %>%
  summarise(Mean=mean(audience_score))

p4a<-ggplot(eda, aes(x=audience_score, color=mpaa_rating_R, fill=mpaa_rating_R))+
  geom_density()+facet_grid(mpaa_rating_R~.)+
  geom_vline(data=mu4, aes(xintercept=Mean),
             linetype="dashed")+
  theme(legend.position = "right", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Distribution")+
  labs(x="Audience Score", y="Density")
```

```{r echo=F, fig.align='center', fig.width=11, fig.height=4 }

kable(eda3)

grid.arrange(p4, p4a, ncol=2)

```

No difference in audience score with an R rating.

#### Oscar Season

```{r echo=F}

eda4<-eda %>%
  group_by(oscar_season) %>%
  summarise(n(), min(audience_score), quantile(audience_score, .25), 
            median(audience_score), mean(audience_score),
            quantile(audience_score, .75), max(audience_score))

names(eda4)<-c("Oscar Season", "Count", "Min", "1st Qu.", "Median", "Mean",
               "3rd Qu.", "Max")
eda4<-mutate(eda4, Skew=ifelse(Mean>Median, "Right", "Left"))

p5<-ggplot(eda, aes(x=factor(oscar_season), y=audience_score, fill=factor(oscar_season)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Box Plot")+
  labs(x="Released during Oscar Season?", y="Audience Score")

mu5<-eda %>%
  group_by(oscar_season) %>%
  summarise(Mean=mean(audience_score))

p5a<-ggplot(eda, aes(x=audience_score, color=oscar_season, fill=oscar_season))+
  geom_density()+facet_grid(oscar_season~.)+
  geom_vline(data=mu5, aes(xintercept=Mean),
             linetype="dashed")+
  theme(legend.position = "right", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Distribution")+
  labs(x="Audience Score", y="Density")

```

```{r echo=F, fig.align='center', fig.width=11, fig.height=4}

kable(eda4)

grid.arrange(p5, p5a, ncol=2)

```

Movies released during Oscar season have a slightly higher average audience score.

#### Summer Season

```{r echo=F}

eda5<-eda %>%
  group_by(summer_season) %>%
  summarise(n(), min(audience_score), quantile(audience_score, .25), 
            median(audience_score), mean(audience_score),
            quantile(audience_score, .75), max(audience_score))

names(eda5)<-c("Summer Release", "Count", "Min", "1st Qu.", "Median", "Mean",
               "3rd Qu.", "Max")
eda5<-mutate(eda5, Skew=ifelse(Mean>Median, "Right", "Left"))

p6<-ggplot(eda, aes(x=factor(summer_season), y=audience_score, fill=factor(summer_season)))+
  geom_boxplot()+
  theme(legend.position = "none", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Box Plot")+
  labs(x="Summer Release?", y="Audience Score")

mu6<-eda %>%
  group_by(summer_season) %>%
  summarise(Mean=mean(audience_score))

p6a<-ggplot(eda, aes(x=audience_score, color=summer_season, fill=summer_season))+
  geom_density()+facet_grid(summer_season~.)+
  geom_vline(data=mu6, aes(xintercept=Mean),
             linetype="dashed")+
  theme(legend.position = "right", axis.text.x = element_text(angle = 60,hjust = 1))+
  ggtitle("Audience Score Distribution")+
  labs(x="Audience Score", y="Density")
```

```{r echo=F, fig.align='center', fig.width=11, fig.height=4}

kable(eda5)

grid.arrange(p6, p6a, ncol=2)

```




* * *

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

